---
title: 'Developing Data Products Course Project: Shiny Application and Reproducible
  Pitch'
author: "YCChen"
date: "2017/4/28"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## My APP: Wave Superposition  

Did you know mulit-single waves can combine as a beautiful wave shape?  
  
This APP will take a tour of this. 
  
We provide 6 kinds of wave with wavelength 0.5 ~ 3 to adjust their amplitude. Final result is superposition of those wave you specify.  
  
Try it yourself to create your own beautiful wave.  

APP: [https://gitycc.shinyapps.io/wave_superposition/](https://gitycc.shinyapps.io/wave_superposition/)  

Shiny Source Code: [https://github.com/GitYCC/Coursera_DataScience_  
DevelopingDataProducts_Project/blob/master/  
WaveSuperposition/app.R](https://github.com/GitYCC/Coursera_DataScience_DevelopingDataProducts_Project/blob/master/WaveSuperposition/app.R)

## Usage

- adjust bars to select amplitude from -10 ~ 10
- wavelength we provided are 0.5, 1, 1.5, 2, 2.5 and 3
- the thick line in plot mean superposition of waves
- fine lines in plot mean wave elements  
     * blue line: wavelength = 0.5
     * orange line: wavelength = 1.0
     * red line: wavelength = 1.5
     * green line: wavelength = 2.0
     * purple line: wavelength = 2.5
     * brown line: wavelength = 3.0

   

## Code (1/2)
load shiny liberary and define ui part
```{r , echo = TRUE}
library(shiny)
ui <- fluidPage(
   titlePanel("Wave Superposition"), # Application title
   sidebarLayout(    # Sidebar 
      sidebarPanel(
         sliderInput("amp_0.5", "Amplitude of wavelength 0.5:",
                    min = -10, max = 10, value = 0),
         sliderInput("amp_1", "Amplitude of wavelength 1:",
                     min = -10, max = 10, value = 10)
           # ... 
      ),
      mainPanel(   # Show a plot 
         plotOutput("allPlot")
      )
   )
)
```

## Code (2/2)
define server part and run APP
```{r , echo = TRUE, eval=FALSE}
server <- function(input, output) {
   output$allPlot <- renderPlot({
      X <- seq(0,6,0.01)
      unit_f <- function(x,wavelength,amp){amp*sin(2*pi*x/wavelength)}
      f <- function(x){ 
          unit_f(x,0.5,input$amp_0.5) + unit_f(x,1.0,input$amp_1) +
          unit_f(x,1.5,input$amp_1.5) + unit_f(x,2.0,input$amp_2) +
          unit_f(x,2.5,input$amp_2.5) + unit_f(x,3.0,input$amp_3) 
      }
      plot(X,f(X),type='l',col='black',lwd=4)
      lines(X,unit_f(X,0.5,input$amp_0.5),type='l',col='blue',lwd=2)
      # ...
   })
}
shinyApp(ui = ui, server = server) # Run the application 
```

